---
title: "R Notebook"
output: html_notebook
---

# Comparison of Rcpp and python implementation.

This requires exfocus installed from this repo: <https://github.com/gtromano/exfocus/tree/main>

```{r}
library(reticulate)
use_python("../../.venv/bin/python")
```

## Importing the libraries

```{python}
from changepoint_online import *
import numpy as np
```

```{r}
library(exfocus)
```

## generating data examples

### gaussian

```{r}
theta0 <- 2
set.seed(42)
Y <- c(rnorm(200, theta0), rnorm(200, theta0 + .2))
```

Pre-change mean unknown

```{r}
res_r <- exfocus(Y, Inf, family = "gaussian")
```

```{python}
detector = Focus(Gaussian())
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
(py$res_py - res_r$stat)[1:10]
```

Pre-change mean known

```{r}
res_r <- exfocus(Y, Inf, family = "gaussian", theta0 = theta0)
```

```{python}
detector = Focus(Gaussian(theta0=r.theta0))
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
cbind(py$res_py, res_r$stat)[1:10,]
```
### Bernoulli

```{r}
theta0 <- .4
set.seed(42)
Y <- c(rbinom(200, prob =  theta0, size = 1), rbinom(200, prob = theta0 + .3, size = 1))
```

Pre-change mean unknown

```{r}
res_r <- exfocus(Y, Inf, family = "bernoulli")
```

```{python}
# testing call to bernoulli
detector = Focus(Bernoulli())
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
cbind(Y, py$res_py, res_r$stat)[1:10, ]
```
### Poisson

Pre-change mean unkown

```{r}
theta0 <- 1
set.seed(42)
Y <- c(rpois(200, theta0), rpois(200, 1.5))
```

Pre-change mean unknown

```{r}
res_r <- exfocus(Y, Inf, family = "poisson")
```

```{python}
detector = Focus(Poisson())
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
cbind(py$res_py, res_r$stat)[1:10, ]
```

Pre-change mean known

```{r}
res_r <- exfocus(Y, Inf, family = "poisson", theta0 = theta0)
```

```{python}
detector = Focus(Poisson(theta0=r.theta0))
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
cbind(py$res_py, res_r$stat)[1:10, ]
```

### Gamma

```{r}
theta0 <- 4
shape <- 1
set.seed(42)
Y <- c(rgamma(200, rate = theta0, shape = shape), rgamma(200, rate = theta0 + 1, shape = shape))
```

Pre-change mean unknown

```{r}
res_r <- exfocus(Y, Inf, family = "gamma", args = list(shape = shape))
```

```{python}
# testing call to gamma
detector = Focus(Gamma(shape=r.shape))
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
cbind(py$res_py, res_r$stat)[1:10, ]
```

Pre-change mean known

```{r}
res_r <- exfocus(Y, Inf, family = "gamma2", args = list(shape = shape), theta0 = 1/theta0)
```

```{python}
detector = Focus(Gamma(theta0=1/r.theta0, shape=r.shape))
res_py = []
for y in r.Y:
    detector.update(y)
    res_py.append(detector.statistic())
```

```{r}
cbind(py$res_py, res_r$stat)[1:10, ]
```

```{python}
detector.ql.ps[10].theta0
```
